\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}

\chapter*{Acknowledgements}

\normalsize
Dr. Sander Wubben and Prof. Dr. Emiel Krahmer for the supervision of this thesis. 

The National Council of Scientific and Technological Development from Brazil (CNPq) and the creators of the Science without Borders program for granting this PhD.

Adriana Baltaretu, Ákos Kádár, Diego Moussallem and Iacer Calixto for collaborating with me and co-authoring some of the research projects developed during this period. 

Prof. Dr. Ivandré Paraboni, who motivated me to look for and start this journey. 

The members of the TiCC language group for reviewing some of my manuscripts and giving insightful comments about them. 

My committee, Prof. Dr. Albert Gatt, Prof. Dr. Antal van den Bosch, Dr. Claire Gardent, Dr. Frank Schilder and Dr. Martijn Goudbeek, for their comments in this thesis.

My mother, Angela Maria Castro Ferreira, my sister, Tatiane Castro Ferreira, my father, Tomé Ferreira Neto, my grandmother, Vicentina Festagallo Castro, my aunt, Ivanir Aparecida Ferreira Pinto Spada, for being my family.

My TiCC colleagues for providing a very nice work environment, specially Eva Verschoor and Lauraine de Lima for all the assistance whenever I needed.

My good friends and colleagues who I met in The Netherlands, specially Katja Helminski (Katjaaaaaaaa*, she will understand when she reads it), Moinuddin M. Haque (Moin), Ayane Santos, René Almeida, Mariana Rachel Dias da Silva (Coração), Mirjam de Haas, Alexandra Sierra, Brenda Szongoth, Evgeny Lavrentjev (Zhenya), Yueqiao Han (Ms. Julie), Adriana Baltaretu, Yevgen Matusevych (Zhenya), Ákos Kádár, Nadine Braun, Priscila Osório Côrtes, Bram Willemsen, Chris Emmery (Metal Chris), Chris van der Lee (Tall Chris), Nanne van Noord, Christine Cook (Chrissy), Laura Capera, Wilma Latuny, Yaser Norouzzadeh Ravari, Yu Gu, Lisa Rombout, Rocsana Bianca and Koel Dutta Chowdhury. Also Fernanda França (Fefa), Jon De Jonge (Jon), Kelly Hessels, Alina Steblovskaya, Dagan Nathaniel Bland and “The master friends” group, which I hope it will become “The friends of the PhD” group: Lívia Riye, Giovana Cremasco, Luis Cristovao, Juliana Almeida, Daniel Guedes, Paula Meira Chinelato (Paulinha), Joana Mattei and Geovana Reis.

My friends from high school and aggregates, Amanda Penna (Mandy), Arthur Grava (Tutu), Bruno Arrabal, Carla Campinas, Caroline Kerestes (Keka), Diogo Gouveia, Flávio Aldana, Gabriele A. de Almeida (Gabs), Juliana Biancheze, Juliana Wes, Kaíssa Nascimento, Mariana Biancheze (Mari), Mariela Ribeiro, Renan Kenji, Renan Rodrigues (Demo), Samira Alvarez Sardella, Vinícius Fernandes (Vini), Vitor Oliveira (Ponto) and the baby members, Rafael Quina Kerestes de Oliveira and Theo Biancheze de Alencar Bravo.

All my friends from Itaú Bank and University of São Paulo (USP), specially Jean Adam Calixto, Márcia Cristina N. Costa, Evaldo Nigro, \linebreak Marcio Coutinho, Juliana Coutinho Rapanelli, Eder Novais, Mirella Hüne, Alessandro Costa, Rafael Ribeiro, Marcus Viudes, Ricardo Rodrigo Santos (Tetra), Lucas Lima, Luan Kendji, Carlos Eduardo Cagna (Picanha) and João Roisin.

In sum, it was the interactions with all of you, in the many good and bad moments during this period, which resulted in this very special chapter in the book of my life. Thank you!