\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}

\chapter*{Acknowledgements}

I would like to thank my main supervisor  Afra Alishahi for providing me with the PhD opportunity 
and NWO for the funding. Thanks also go to my second supervisor Grzegorz Chrupała for joining the project
and providing his valuable efforts in improving the thesis. 
They started me off on my trajectory as a researcher and will have a long lasting 
impact on my further development.
Eric Postma, my promotor, oversaw my trajectory and provided crucial input especially at the finishing stages
for which I'm grateful. 

Thanks go to the Integrating Vision and Language COST Action unit for providing me with the
"Short Term Scientific Mission grant IC1307" and for Tartu University for hosting. 
Tambet Matiisen was an exceptional host and I hope that we have an entangled future ahead of us.
Further, thanks go to the organizers of the "Integrating Vision and Language Summer School" for 
awarding me the "Trainee grant" twice in a row and to the organizers of the "LxMLS Summer School"
for awarding me their trainee grant. 

I had a wonderful time working for Microsoft Research Montreal during my five months internship
and I am very grateful for the opportunity and for the mentorship my colleagues provided.

I would also like to thank my collaborators Desmond Elliot and Marc-Alexandre Côté for their contributions
to my thesis. It was such a pleasure working together and I hope we will have the opportunity to continue in the
future despite the current dystopia created around "intellectual property rights".


I couldn't have pushed through my last year without getting pumped with endorphins 
by the wonderful quintet of Strawberry, Lychee, Lobsterwoman, ElfHorse and FoxFace . 
Time of my life girls , I will never forget you.  
Looking forward to seeing you in Toronto FoxFace! Given how horrible you are I cannot believe
it that you have so many lines dedicated to you in this thesis. Love you!
My shrink Ibrahim Akkaya also did a lot of work on keeping brain chemistry at bay; 
I would have probably quit after around 15-18 months if it wasn't for him. 

The future's uncertain and the end is always near. Let it roll, baby, roll!



%So yeah, I am grateful that my supervision team pushed through all this madness and I would especially like to thank my promotor Eric Postma
%for helping me chill the f*ck out when things were very bad. I am more grateful though for the people who helped me get through this
%incredibly backwards world of the PhD. 



%Big shoutout goes to the sociologist 
%crew from Trento University who spend half their time in Tilburg. Gabriele and Francesca my old friends, I will always love you. The younger 
%ones too, Lucilla, Laura, of course Claudia and Giovanni the greatest junkie of all. Due to your beautiful presence in my life I will always think
%f marxist Italian shitheads as the most beautiful people on the planet. 

%Completing this thesis is an unpleasant experience.
%You see, a limitation of mine is the inability to view stuff around me as stationary
%things, rather, I tend to see everything as processeses.
%Never managed to grasp the idea that there are moments that have passed,
%that there is the ``right now'' and that there is a future to look forward to.
%The ``starting'' and ``finishing'' of this work is imposed upon me
%by convention and to be honest it was very challenging.
%%My projects started from the desire to do something and were turned into
%%finished papers, because that is how we do things.
%Anyhow, now you are here reading the start of the end of this chapter of mine
%so I should try to put some nice words here. Don't get me wrong,
%I do understand that now is the time to move on to first thank my supervisors
%then my family followed by my colleagues and then finish up with perhaps a bit
%cryptic and personal messages to some special others. My problem is that I
%feel like I was always very open about my appretiation, greatfulness, criticism
%and disregard. I am not sure what there is to say for me here that I have not
%said to You in person. Furthermore, I handle formalities incredibly poorly.
%To my supervisors I
%would say that they did a much better job than a lot of others. Authoring
%a message to my family here seems a bit useless to me.
%I also hope that the people I love are not going
%go through the struggle of reading any of this stuff. I'm clearly not a
%great writer and the topic of my thesis is very tangental to the lives of the
%vast majority. As it is obvious by now I personally fail to see
%the epic significance of this \LaTeX document and I will continue anyway to
%indulge in my relationships with the wonderful human beings who share their
%lives with me.
